---
title: "Data 609 - Module2"
author: "Amit Kapoor"
date: "9/10/2021"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
  html_document:
    fig_width: 15
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Ex. 1
Write down Newton's formula for finding the minimum of $f(x) =  (3x^{4} -  4x^{3}) / 12$ in the range [-10, 10]. Then, implement it in R.

**Solution:**

$f(x) =  (3x^{4} -  4x^{3}) / 12$

$f'(x) = \frac{12x^{3} - 12x^2}{12} =  x^3 - x^2$

$f''(x) = 3x^2 - 2x$

Using Newton's formula

$x_{k  + 1} =  x_{k} -  \frac{x_k^{3} - x_k^{2}}{3x_k^{2} - 2x_k}$

```{r}

f <- function(x) {
  x - (x^3 - x^2)/(3 * x^2 - 2*x)
}

fetch_values <- function(X_0) {
  val <- c()
  for(i in 1:10) {
    if(i==1) {
      val[[i]] <- f(X_0)
    } else {
      val[[i]] <- f(val[[i-1]])
    }
  }
  print(val)
}

```


```{r}
fetch_values(3)
```


```{r}
fetch_values(1.5)
```


It looks like the minimum is at 1 since given multiple starting values, it converges to 1.




# Ex. 2
Explore optimize() in R and try to solve the previous problem.

**Solution:**

```{r}
f <- function(x) {
  (3*x^4 -  4*x^3)/12
}

x_min <- optimize(f, c(-10, 10))
x_min
```

Here we can confirm that minimum is at 1 and value of given function is -0.0833.


# Ex. 3 
Use any optimization algorithm to find the minimum of $f(x,y) =  (x-1)^{2} + 100 (y -  x^{2}) ^{2}$ in the domain $-10  \leq x,y  \leq 10.$ Discuss any issues concerning the optimization process.

**Solution:**

Ww will use Newton's method for optimization of multivariate function. 

The equation is

$x_{t + 1} =  x_{t} -  H^{-1}\bigtriangledown f(x,y)$

where $\bigtriangledown f$ is the gradient vector and $H$ is Hessian matrix

Starting value.

$x_{0} = \begin{bmatrix}x  \\y \end{bmatrix} =\begin{bmatrix}0  \\0 \end{bmatrix}$ 

$\bigtriangledown f(x,y) =  \begin{bmatrix}(x-1)^{2}  \\100 (y -  x^{2}) ^{2} \end{bmatrix}$

We will start by finding Hessian matrix.

$f_{xx} = 1200 x^{2} + 2 - 400y$

$f_{xy} = -400x$

$f_{yy} = 200$

$f_{yx} = -400x$

$H = \bigtriangledown^2 f(x,y) =  \begin{bmatrix}2 & -0  \\-0 & 200 \end{bmatrix}$ 

$H^{-1}  = \begin{bmatrix}0.5 & 0  \\0 & 0.005 \end{bmatrix}$ 

With t = 0 $x_{1} =  x_{0} -  H^{-1}\bigtriangledown f(x,y)$

$x_{1} =  \begin{bmatrix}0  \\0 \end{bmatrix} - \begin{bmatrix}0.5 & 0  \\0 & 0.005 \end{bmatrix}\begin{bmatrix}1  \\0 \end{bmatrix}$

$x_{1} =  \begin{bmatrix}0  \\0 \end{bmatrix} - \begin{bmatrix}0.5  \\0 \end{bmatrix}$

$x_{1} =  \begin{bmatrix}-0.5  \\0 \end{bmatrix}$

Given this, we can say f(x,y) will converge into a single point as we go through further values of t.


# Ex. 4
Explore the optimr package for R and try to solve the previous problem.

**Solution:**

```{r lib, include=FALSE}
# Libraries
library(optimr)
```

```{r}

f_xy <- function(param) {
  (param[1] - 1)^2 + 100 * (param[2] - param[1]^2)^2
}

options(scipen = 999)
optimr(c(-10,10), f_xy, method = "L-BFGS-B")
```













