---
title: "Data 609 - Module7"
author: "Amit Kapoor"
date: "11/23/2021"
output:
  html_document:
    fig_width: 15
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries}
# Libraries
library(e1071)
library(caret)
```


# Ex.1
Use the svm() algorithm of the `e1071` package to carry out the support vector machine for the `PlantGrowth` data set. Then, discuss the number of support vectors/samples. [Install the e1071 package in R if needed.]

**Solution**

```{r pg-data}
# PlantGrowth datset
data(PlantGrowth)
head(PlantGrowth)
```
```{r plant-svm}
svm_pg <- svm(group ~ ., data = PlantGrowth)
summary(svm_pg)
```

It comes out that there are 29 support vectors out of 30 samples that are closer to the hyperplane and influence the orientation and position of the hyperplane. There are 3 classes.




# Ex.2
Do a similar SVM analysis as that in the previous question using the `iris` data set. Discuss the number of support vectors/samples.

**Solution**

```{r iris-data}
# iris datset
data("iris")
head(iris)
```

```{r iris-svm}
svm_iris <- svm(Species ~ ., data = iris)
summary(svm_iris)
```

There are 51 support vectors and 3 classes in this case out of 150 samples in the dataset. In this case there is very less data for support vectors as compared to `PlantGrowth` dataset.

# Ex.3
Use the iris data set (or any other data set) to select 80% of the samples for the training svm(), then use the rest 20% for validation. Discuss your results.

**Solution**

```{r iris-train}
set.seed(609)

# partitioning for training and validation
partition <- createDataPartition(iris$Species, p=0.80, list = FALSE)
training <- iris[partition,]
validation <- iris[-partition,]


svm_train_iris <- svm(Species ~ ., data = training)
summary(svm_train_iris)

```


```{r iris-pred}
pred_valid <- predict(svm_train_iris, validation)
confusionMatrix(validation$Species, pred_valid)
```


The accuracy of the model has comeout as 96.67%.Here the training model shows 45 support vectors.







