---
title: "Data 609 - Module2"
author: "Amit Kapoor"
date: "9/10/2021"
output:
  html_document:
    fig_width: 15
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Ex. 1
Show $x^{2} + exp(x) + 2 x^{4} + 1$ is convex.

**Solution:**

A function is considered convex if
$f( \alpha x +  \beta y)  \leq   \alpha f(x) +  \beta f(y)$, $\forall x,y \in \Omega$, $\alpha \ge 0, \beta \ge 0$, $\alpha + \beta = 1$

$(\alpha x +  \beta y)^{2} + exp(\alpha x +  \beta y) + 2 (\alpha x +  \beta y)^{4} + 1  \leq   \alpha (x^{2} + exp(x) + 2 x^{4} + 1) +  \beta (y^{2} + exp(y) + 2 y^{4} + 1)$

$(\alpha^{2} x^{2} +  \beta^{2} y^{2} + 2 \alpha \beta x y ) + exp(\alpha x +  \beta y) + 2 (\alpha x +  \beta y)^{4} + 1  \leq   \alpha x^{2} + \alpha exp(x) + 2 \alpha x^{4} + \alpha +  \beta y^{2} + \beta exp(y) + 2 \beta y^{4} + \beta$

Using $\alpha + \beta = 1$ we rewrite as 

$(\alpha^{2} x^{2} +  \beta^{2} y^{2} + 2 \alpha \beta x y ) + exp(\alpha x +  \beta y) + 2 (\alpha x +  \beta y)^{4} \leq   \alpha x^{2} + \alpha exp(x) + 2 \alpha x^{4} +  \beta y^{2} + \beta exp(y) + 2 \beta y^{4}$

$2 \alpha  x^{4} +  \alpha  x^{2} +  \alpha exp(x) + 2 \beta  y^{4} +  \beta  y^{2} +  \beta exp(y) - (\alpha x +  \beta y)^{2} - exp(\alpha x -  \beta y) - 2 (\alpha x +  \beta y)^{4} \geq 0$

The inequality is always true and hence $x^{2} + exp(x) + 2 x^{4} + 1$ is convex.

# Ex. 2
Show that the mean of the exponential distribution 

$p(x) = \begin{cases} \lambda  e^{- \lambda x},  & x  \geq  0 ( \lambda  > 0),\\0, & x  <  0\end{cases}$ 

is $\mu = \frac{1}{ \lambda }$ and its variance is $\sigma ^{2} =  \frac{1}{  \lambda ^{2} }$.

**Solution:**

To find the mean we know that $\mu = E[X] =\int xp(x)dx$

$\mu = \int_{-\infty}^0 x*0dx +  \int_0^ \infty  x \lambda  e^{- \lambda x} dx$

Using integration by parts $\int u v dx = u  \int v dx -  \int  u' (\int v dx) dx$

replacing $u=x, v=\lambda  e^{- \lambda x}$ we will have

$\mu =[-x e^{- \lambda x}]_0^ \infty + \int_0^ \infty e^{- \lambda x} dx$

$\mu =[-x e^{- \lambda x}]_0^ \infty + [ -\frac{1}{ \lambda } e^{- \lambda x}]_0^ \infty$

$\mu = (0 - 0) + (0 + \frac{1}{ \lambda })$ = $\frac{1}{ \lambda }$

To find the variance of the given exponential distribution we know that $\sigma ^{2} = Var[X] = E[ X^{2}] - E[X] ^{2}$

$E[ X^{2}] = \int_0^ \infty   x^{2} \lambda  e^{- \lambda x}$ which could again be solved by using integration by parts

$E[ X^{2}] = [-x^{2} e^{- \lambda x}]_0^ \infty + \int_0^ \infty 2xe^{- \lambda x} dx$ 

$E[ X^{2}] = [-x^{2} e^{- \lambda x}]_0^ \infty +[- \frac{2}{ \lambda } x e^{- \lambda x} dx]_0^ \infty +   \frac{2}{ \lambda } \int_0^ \infty e^{- \lambda x} dx$ 

using integration by parts again

$E[ X^{2}] = [-x^{2} e^{- \lambda x}]_0^ \infty +[- \frac{2}{ \lambda } x e^{- \lambda x} dx]_0^ \infty +  \frac{2}{ \lambda } [- \frac{1}{ \lambda } x e^{- \lambda x} dx]_0^ \infty$

$(\frac{2}{ \lambda })( \frac{1}{ \lambda })$ = $\frac{2}{  \lambda ^{2}}$

$E[ X^{2}] - E[X] ^{2} = \frac{2}{\lambda ^{2}} - (\frac{1}{ \lambda }) ^{2} = \frac{2}{\lambda ^{2}} - \frac{1}{\lambda ^{2}} = \frac{1}{  \lambda ^{2}}$







# Ex. 3
It is estimated that there is a typo in every 250 data entries in a database, assuming the number of typos can obey the Poisson distribution. For a given 1000 data entries, what is the probability of exactly 4 typos? What is the probability of no typo at all? Use R to draw 1000 samples with $\lambda = 4$ and show their histogram.



**Solution:**

```{r}
lambda <- 1000 * (1/250)
x <- 4 # 4 typos
p_four_typo <- (lambda ^ x * exp(-lambda)) / factorial(x)
p_four_typo
```

The probability of exactly 4 typos is 19.54%.

```{r}
lambda <- 4
x <- 0 # no typo
p_zero_typo <- (lambda ^ x * exp(-lambda)) / factorial(x)
p_zero_typo
```

The probability of no typos at all is 1.83%.


Now let's confirm our results above using built in R function.

```{r}
# probability of exactly 4 typos
p1 <- dpois(4, lambda = 4)
p1
```

```{r}
# probability of no typo
q2 <- dpois(0, lambda = 4)
q2
```

```{r}
events = 1:1000
lambda = 4
poisson <- rpois(1000, lambda)
df <- data.frame(events, poisson)

hist(df$poisson, main = "1000 Samples with lambda = 4")
```
