---
title: "Data 609 - Module6"
author: "Amit Kapoor"
date: "10/22/2021"
output:
  html_document:
    fig_width: 15
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
# Libraries
library(class)
library(cluster)
```

# Ex. 1
Use a data set such as the `PlantGrowth` in R to calculate three different distance metrics and discuss the results.

**Solution**

We will use here `PlantGrowth` dataset in R and dist() function that gives the distance matrix using specified distance to compute the distances between the rows of distance matrix. Here the methods considered are `manhattan`, `euclidean` and `canberra`.

```{r man}
plants_man <- dist(PlantGrowth, method = "manhattan")
as.matrix(plants_man)
```


```{r euc}
plants_euc <- dist(PlantGrowth, method = "euclidean")
as.matrix(plants_euc)
```



```{r can}
plants_can <- dist(PlantGrowth, method = "canberra")
as.matrix(plants_can)
```




# Ex. 2
Now use a higher-dimensional data set `mtcars`, try the same three distance metrics in the previous question and discuss the results.

**Solution**

Now we will have `mtcars` dataset.

```{r mtcars}
head(mtcars)
```

We will take transpose of mtcars dataset and apply the same 3 distance methods.

```{r t-mtcars}
head(t(mtcars))
```




```{r mtcars-man}
mtcars_man <- dist(t(mtcars), method = "manhattan")
as.matrix(mtcars_man)
```



```{r mtcars-euc}
mtcars_euc <- dist(t(mtcars), method = "euclidian")
as.matrix(mtcars_euc)
```



```{r mtcars-can}
mtcars_can <- dist(t(mtcars), method = "canberra")
as.matrix(mtcars_can)
```




# Ex. 3
Use the built-in data set `mtcars` to carry out hierarchy clustering using two different distance metrics and compare if they get the same results. Discuss the results.

**Solution**



```{r summ-mtcars}
summary(mtcars)
```


```{r euc-clus}
euc_clus <- hclust(dist(mtcars, method = "euclidian"))
plot(euc_clus)
```


```{r euc-ct}
ct1 <- cutree(euc_clus, k = 3)
table(ct1)
```


```{r man-clus}
man_clus <- hclust(dist(mtcars, method = "manhattan"))
plot(man_clus)
```


```{r man-ct}
ct2 <- cutree(man_clus, k = 3)
table(ct2)
```


We see above both the methods give higher elements to first cluster but second and third clusters are more evenly distributed in euclidian. 



# Ex. 4
Load the well-known Fisher’s `iris` flower data set that consists of 150 samples for 3 species (50 samples each species). The four measures or features are the lengths and widths of sepals and petals. Use the kNN clustering to analyze this iris data set by selecting 120 samples for training and 30 samples for testing.


**Solution**



```{r iris}
# iris dataset
head(iris)
```


```{r iris-knn}
set.seed(609)

# to split
split <- sample(nrow(iris), nrow(iris)*0.80)

# feature variables
train <- iris[split, -5] #120 rows
test <- iris[-split, -5] #30 rows

# target variable
train_trgt <- iris[split, 5] #120 rows
test_trgt <- iris[-split, 5] #30 rows

# knn
knn <- knn(train, test, cl=train_trgt, k=5)

# contingency table
knn_tbl <- table(knn, test_trgt)

# knn accuracy
acc_knn <- sum(diag(knn_tbl)) / sum(knn_tbl)
acc_knn
```




# Ex. 5
Use the `iris` data set to carry out k-means clustering. Compare the results to the actual classes and estimate the clustering accuracy.


**Solution**



```{r}
set.seed(609)

iris_kmeans <- kmeans(iris[,-5], centers = 3)
iris_kmeans
```


```{r}
# contingency table
iris_clus <- iris_kmeans$cluster
table(iris_clus)
```



```{r clusplot}
clusplot(iris[,-5], iris_clus)
```


```{r}
# kmeans accuracy
tbl_kmeans <- table(iris$Species, iris_clus)
acc_kmeans <- sum(diag(tbl_kmeans)) / sum(tbl_kmeans)
acc_kmeans
```













